---
title: "Business Intelligence Lab Submission Markdown"
author: "<Specify your group name here>"
date: "<Specify the date when you submitted the lab>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+---------------------------------------------+
| **Student ID Numbers and Names of Group Members** | 1.  133996 - B - Trevor Ngugi               |
|                                                   |                                             |
|                                                   | 2.  133996 - B - Immaculate Haayo           |
|                                                   |                                             |
|                                                   | 3.  135859 - B - Pauline Wairimu            |
|                                                   |                                             |
|                                                   | 4.  127701 - B - Clarice Gitonga            |
|                                                   |                                             |
|                                                   | 5.  126761 - B - Virginia Wanjiru           |
+---------------------------------------------------+---------------------------------------------+
| **GitHub Classroom Group Name**                   | Champions                                   |
+---------------------------------------------------+---------------------------------------------+
| **Course Code**                                   | BBT4206                                     |
+---------------------------------------------------+---------------------------------------------+
| **Course Name**                                   | Business Intelligence II                    |
+---------------------------------------------------+---------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology |
+---------------------------------------------------+---------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023  |
+---------------------------------------------------+---------------------------------------------+

# Setup Chunk

We start by installing all the required packages

```{r Install Packages, echo=TRUE, message=FALSE, warning=FALSE}
## formatR - Required to format R code in the markdown ----

# Installed  renv package:
if (!is.element("renv", installed.packages()[, 1])) {
  install.packages("renv", dependencies = TRUE)
}
require("renv")
```

------------------------------------------------------------------------

**Note:** the following "*KnitR*" options have been set as the defaults in this markdown:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy.opts = list(width.cutoff = 80), tidy = TRUE)`.

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
	eval = TRUE,
	echo = TRUE,
	warning = FALSE,
	collapse = FALSE,
	tidy = TRUE
)
```

------------------------------------------------------------------------

**Note:** the following "*R Markdown*" options have been set as the defaults in this markdown:

> output:\
> \
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default\
> \
> editor_options:\
> chunk_output_type: console

# Loading the Student Performance Dataset

StudentPerformanceDataset is then loaded. The dataset and its metadata are available here: <https://drive.google.com/drive/folders/1-BGEhfOwquXF6KKXwcvrx7WuZXuqmW9q?usp=sharing>

```{r Load Dataset}

## We loaded the dataset from a csv file

if (!is.element("readr", installed.packages()[, 1])) {
  install.packages("readr", dependencies = TRUE,
                   repos="https://cloud.r-project.org")
}
require("readr")



student_performance_dataset <-
  readr::read_csv(
    "C:/Users/Virginia/OneDrive - Strathmore University/Desktop/Business Intelligence/BBT4206-R-Lab2of15-ExploratoryDataAnalysis-champions/data/StudentPerformanceDataset.csv", 
    col_types =
      readr::cols(
        class_group =
          readr::col_factor(levels = c("A", "B", "C")),
        gender = readr::col_factor(levels = c("1", "0")),
        YOB = readr::col_date(format = "%Y"),
        regret_choosing_bi =
          readr::col_factor(levels = c("1", "0")),
        drop_bi_now =
          readr::col_factor(levels = c("1", "0")),
        motivator =
          readr::col_factor(levels = c("1", "0")),
        read_content_before_lecture =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        anticipate_test_questions =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        answer_rhetorical_questions =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        find_terms_I_do_not_know =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        copy_new_terms_in_reading_notebook =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        take_quizzes_and_use_results =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        reorganise_course_outline =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        write_down_important_points =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        space_out_revision =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        studying_in_study_group =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        schedule_appointments =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        goal_oriented =
          readr::col_factor(levels =
                              c("1", "0")),
        spaced_repetition =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        testing_and_active_recall =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        interleaving =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        categorizing =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        retrospective_timetable =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        cornell_notes =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        sq3r = readr::col_factor(levels =
                                   c("1", "2", "3", "4")),
        commute = readr::col_factor(levels =
                                      c("1", "2",
                                        "3", "4")),
        study_time = readr::col_factor(levels =
                                         c("1", "2",
                                           "3", "4")),
        repeats_since_Y1 = readr::col_integer(),
        paid_tuition = readr::col_factor(levels =
                                           c("0", "1")),
        free_tuition = readr::col_factor(levels =
                                           c("0", "1")),
        extra_curricular = readr::col_factor(levels =
                                               c("0", "1")),
        sports_extra_curricular =
          readr::col_factor(levels = c("0", "1")),
        exercise_per_week = readr::col_factor(levels =
                                                c("0", "1",
                                                  "2",
                                                  "3")),
        meditate = readr::col_factor(levels =
                                       c("0", "1",
                                         "2", "3")),
        pray = readr::col_factor(levels =
                                   c("0", "1",
                                     "2", "3")),
        internet = readr::col_factor(levels =
                                       c("0", "1")),
        laptop = readr::col_factor(levels = c("0", "1")),
        family_relationships =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        friendships = readr::col_factor(levels =
                                          c("1", "2", "3",
                                            "4", "5")),
        romantic_relationships =
          readr::col_factor(levels =
                              c("0", "1", "2", "3", "4")),
        spiritual_wellnes =
          readr::col_factor(levels = c("1", "2", "3",
                                       "4", "5")),
        financial_wellness =
          readr::col_factor(levels = c("1", "2", "3",
                                       "4", "5")),
        health = readr::col_factor(levels = c("1", "2",
                                              "3", "4",
                                              "5")),
        day_out = readr::col_factor(levels = c("0", "1",
                                               "2", "3")),
        night_out = readr::col_factor(levels = c("0",
                                                 "1", "2",
                                                 "3")),
        alcohol_or_narcotics =
          readr::col_factor(levels = c("0", "1", "2", "3")),
        mentor = readr::col_factor(levels = c("0", "1")),
        mentor_meetings = readr::col_factor(levels =
                                              c("0", "1",
                                                "2", "3")),
        `Attendance Waiver Granted: 1 = Yes, 0 = No` =
          readr::col_factor(levels = c("0", "1")),
        GRADE = readr::col_factor(levels =
                                    c("A", "B", "C", "D",
                                      "E"))),
    locale = readr::locale())

```

## Description of the Dataset

We then display the number of observations and number of variables. We have 101 observations and 100 variables in our dataset.

```{code to display the datasets dimensions}
dim(student_performance_dataset)
```

In this step, we identify the data types in the datasets which will help us in determining which visualization types are most appropriate to be used.

## Identifying Data Types 

```{r Your Fifth Code Chunk}
sapply(student_performance_dataset,class)

```

## Measures of Frequency

In the code illustrated below will be checking the frequency of students who take down important points, read content before the lecture and schedule appointments to consult.

```{r Your Sixth Code Chunk}
student_density_freq <- student_performance_dataset$write_down_important_points
cbind(frequency = table(student_density_freq),
      percentage = prop.table(table(student_density_freq)) * 100)

student_density_freq <- student_performance_dataset$read_content_before_lecture
cbind(frequency = table(student_density_freq),
      percentage = prop.table(table(student_density_freq)) * 100)

student_density_freq <- student_performance_dataset$schedule_appointments
cbind(frequency = table(student_density_freq),
      percentage = prop.table(table(student_density_freq)) * 100)
```

## Measure of Distribution

```{r Your Seventh Code Chunk}
summary(student_performance_dataset)


sapply(student_performance_dataset[, 99], sd)
sapply(student_performance_dataset[, 92], sd)
sapply(student_performance_dataset[, 91], sd)
sapply(student_performance_dataset[, 90], sd)
sapply(student_performance_dataset[, c(99,92,91)], sd)



```

## \<You Can Have Another Sub-Title Here if you wish\>

```{r Your Eighth Code Chunk}
# Fill this with other R related code that will be executed when the R markdown
```

**etc.** as per the lab submission requirements. Be neat and communicate in a clear and logical manner.
