---
title: "Business Intelligence Lab Submission Markdown"
author: "<Champions>"
date: "<22/09/2023>"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+---------------------------------------------+
| **Student ID Numbers and Names of Group Members** | 1.  133996 - B - Trevor Ngugi               |
|                                                   |                                             |
|                                                   | 2.  133996 - B - Immaculate Haayo           |
|                                                   |                                             |
|                                                   | 3.  135859 - B - Pauline Wairimu            |
|                                                   |                                             |
|                                                   | 4.  127701 - B - Clarice Gitonga            |
|                                                   |                                             |
|                                                   | 5.  126761 - B - Virginia Wanjiru           |
+---------------------------------------------------+---------------------------------------------+
| **GitHub Classroom Group Name**                   | Champions                                   |
+---------------------------------------------------+---------------------------------------------+
| **Course Code**                                   | BBT4206                                     |
+---------------------------------------------------+---------------------------------------------+
| **Course Name**                                   | Business Intelligence II                    |
+---------------------------------------------------+---------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology |
+---------------------------------------------------+---------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023  |
+---------------------------------------------------+---------------------------------------------+

# Setup Chunk

We start by installing all the required packages

```{r Install Packages, echo=TRUE, message=FALSE, warning=FALSE}
## formatR - Required to format R code in the markdown ----

# Installed  renv package:
if (!is.element("renv", installed.packages()[, 1])) {
  install.packages("renv", dependencies = TRUE)
}
require("renv")
```

------------------------------------------------------------------------

**Note:** the following "*KnitR*" options have been set as the defaults in this markdown:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy.opts = list(width.cutoff = 80), tidy = TRUE)`.

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
	eval = TRUE,
	echo = TRUE,
	warning = FALSE,
	collapse = FALSE,
	tidy = TRUE
)
```

------------------------------------------------------------------------

**Note:** the following "*R Markdown*" options have been set as the defaults in this markdown:

> output:\
> \
> github_document:\
> toc: yes\
> toc_depth: 4\
> fig_width: 6\
> fig_height: 4\
> df_print: default\
> \
> editor_options:\
> chunk_output_type: console

# Loading the Student Performance Dataset

StudentPerformanceDataset is then loaded. The dataset and its metadata are available here: <https://drive.google.com/drive/folders/1-BGEhfOwquXF6KKXwcvrx7WuZXuqmW9q?usp=sharing>

```{r Load Dataset}

## We loaded the dataset from a csv file

if (!is.element("readr", installed.packages()[, 1])) {
  install.packages("readr", dependencies = TRUE,
                   repos="https://cloud.r-project.org")
}
require("readr")



student_performance_dataset <-
  readr::read_csv(
    "C:/Users/Pauline Wang'ombe/Downloads/4.2/BUSINESS INTELLIGENCE 2/BBT4206-R-Lab2of15-ExploratoryDataAnalysis-champions/data/StudentPerformanceDataset.csv", 
    col_types =
      readr::cols(
        class_group =
          readr::col_factor(levels = c("A", "B", "C")),
        gender = readr::col_factor(levels = c("1", "0")),
        YOB = readr::col_date(format = "%Y"),
        regret_choosing_bi =
          readr::col_factor(levels = c("1", "0")),
        drop_bi_now =
          readr::col_factor(levels = c("1", "0")),
        motivator =
          readr::col_factor(levels = c("1", "0")),
        read_content_before_lecture =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        anticipate_test_questions =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        answer_rhetorical_questions =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        find_terms_I_do_not_know =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        copy_new_terms_in_reading_notebook =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        take_quizzes_and_use_results =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        reorganise_course_outline =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        write_down_important_points =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        space_out_revision =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        studying_in_study_group =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        schedule_appointments =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        goal_oriented =
          readr::col_factor(levels =
                              c("1", "0")),
        spaced_repetition =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        testing_and_active_recall =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        interleaving =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        categorizing =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        retrospective_timetable =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        cornell_notes =
          readr::col_factor(levels =
                              c("1", "2", "3", "4")),
        sq3r = readr::col_factor(levels =
                                   c("1", "2", "3", "4")),
        commute = readr::col_factor(levels =
                                      c("1", "2",
                                        "3", "4")),
        study_time = readr::col_factor(levels =
                                         c("1", "2",
                                           "3", "4")),
        repeats_since_Y1 = readr::col_integer(),
        paid_tuition = readr::col_factor(levels =
                                           c("0", "1")),
        free_tuition = readr::col_factor(levels =
                                           c("0", "1")),
        extra_curricular = readr::col_factor(levels =
                                               c("0", "1")),
        sports_extra_curricular =
          readr::col_factor(levels = c("0", "1")),
        exercise_per_week = readr::col_factor(levels =
                                                c("0", "1",
                                                  "2",
                                                  "3")),
        meditate = readr::col_factor(levels =
                                       c("0", "1",
                                         "2", "3")),
        pray = readr::col_factor(levels =
                                   c("0", "1",
                                     "2", "3")),
        internet = readr::col_factor(levels =
                                       c("0", "1")),
        laptop = readr::col_factor(levels = c("0", "1")),
        family_relationships =
          readr::col_factor(levels =
                              c("1", "2", "3", "4", "5")),
        friendships = readr::col_factor(levels =
                                          c("1", "2", "3",
                                            "4", "5")),
        romantic_relationships =
          readr::col_factor(levels =
                              c("0", "1", "2", "3", "4")),
        spiritual_wellnes =
          readr::col_factor(levels = c("1", "2", "3",
                                       "4", "5")),
        financial_wellness =
          readr::col_factor(levels = c("1", "2", "3",
                                       "4", "5")),
        health = readr::col_factor(levels = c("1", "2",
                                              "3", "4",
                                              "5")),
        day_out = readr::col_factor(levels = c("0", "1",
                                               "2", "3")),
        night_out = readr::col_factor(levels = c("0",
                                                 "1", "2",
                                                 "3")),
        alcohol_or_narcotics =
          readr::col_factor(levels = c("0", "1", "2", "3")),
        mentor = readr::col_factor(levels = c("0", "1")),
        mentor_meetings = readr::col_factor(levels =
                                              c("0", "1",
                                                "2", "3")),
        `Attendance Waiver Granted: 1 = Yes, 0 = No` =
          readr::col_factor(levels = c("0", "1")),
        GRADE = readr::col_factor(levels =
                                    c("A", "B", "C", "D",
                                      "E"))),
    locale = readr::locale())

```

## Description of the Dataset

We then display the number of observations and number of variables. We have 101 observations and 100 variables in our dataset.

```{code to display the datasets dimensions}
dim(student_performance_dataset)
```


## Identifying Data Types 
In this step, we identify the data types in the datasets which will help us in determining which visualization types are most appropriate to be used.

```{r Your Fifth Code Chunk}
sapply(student_performance_dataset, class)

```

## Measures of Frequency

In the code illustrated below will be checking the frequency of students who take down important points, read content before the lecture and schedule appointments to consult.

```{r Your Sixth Code Chunk}
student_density_freq <- student_performance_dataset$write_down_important_points
cbind(frequency = table(student_density_freq),
      percentage = prop.table(table(student_density_freq)) * 100)

student_density_freq <- student_performance_dataset$read_content_before_lecture
cbind(frequency = table(student_density_freq),
      percentage = prop.table(table(student_density_freq)) * 100)

student_density_freq <- student_performance_dataset$schedule_appointments
cbind(frequency = table(student_density_freq),
      percentage = prop.table(table(student_density_freq)) * 100)
```

## Measure of Distribution
In the code below we are measuring the distribution of the data for each variable


```{r Your Seventh Code Chunk}
summary(student_performance_dataset)

 




```

## Measure the standard deviation of each variable
We chose the variables below because they are numeric.
```{r Your Eighth Code Chunk}
sapply(student_performance_dataset[, c(97,78,79,81,82,86,91,92,96,99)], sd)

```

## Measure the variance of each variable
We chose the variables below because they are numeric.
```{r Your Ninth Code Chunk}
sapply(student_performance_dataset[, c(97,78,79,81,82,86,91,92,96,99)], var)

```

## Measure the kurtosis of each variable
The first code is used to install the e1071 package for kurtosis and the second code informs us how often the outliers occur in the results
```{r Your Tenth Code Chunk}
if (!is.element("e1071", installed.packages()[, 1])) {
  install.packages("e1071", dependencies = TRUE)
}
require("e1071")

sapply(student_performance_dataset[,c(97,78,79,81,82,86,91,92,96,99)], kurtosis, type = 2)

```

## Measure the covariance between variables
Most of our variables are positive hence the covariance is positive which means that they are moving in the same direction
```{r Your Eleventh Code Chunk}
student_performance_dataset_cov <- cov(student_performance_dataset[,c(97,78,79,81,82,86,91,92,96,99),])
View(student_performance_dataset_cov)

```

## Measure the correlation between variables
Most of our values are 0.xxx positive or negative which implies little or relationship between the variables

```{r Your Twelveth Code Chunk}
student_performance_dataset_cor <- cor(student_performance_dataset[, c(97,78,79,81,82,86,91,92,96,99),])
View(student_performance_dataset_cor)

```

## Perform ANOVA on the "student_performance_dataset" dataset
One-Way ANOVA can be used to test the effect of studying_in_study_group variable on TOTAL
```{r Your Thirteenth Code Chunk}
student_performance_dataset_one_way_anova <- aov(`TOTAL = Coursework TOTAL + EXAM (100%)`   ~ studying_in_study_group, data = student_performance_dataset)
summary(student_performance_dataset_one_way_anova)

```
 
We can also have a situation where the TOTAL depends not only on space_out_revision but also on the studying_in_study_group variables. A two-way ANOVA can be used to confirm this.
```{r Your Fourteenth Code Chunk}
student_performance_dataset_additive_two_way_anova <- aov(`TOTAL = Coursework TOTAL + EXAM (100%)` ~ studying_in_study_group + space_out_revision, # nolint
                                           data = student_performance_dataset)
summary(student_performance_dataset_additive_two_way_anova)

```

Specifying a (+) between the two independent variables (studying_in_study_group + space_out_revision)implies that they have an additive effect
```{r Your Fifteenth Code Chunk}
student_performance_dataset_interactive_two_way_anova <- aov(`TOTAL = Coursework TOTAL + EXAM (100%)` ~ studying_in_study_group * space_out_revision, # nolint
                                              data = student_performance_dataset)
summary(student_performance_dataset_interactive_two_way_anova)

```

## Create Box and Whisker Plots for each Numeric attribute
We used Box and Whisker Plots to show whether a distribution is skewed or not and if there are outliers
```{r Your Sixteenth Code Chunk}
boxplot(student_performance_dataset[, 96], main = names(student_performance_dataset)[96])

```

## Create Bar Plots for each Categorical Attribute
We used a Bar Chart to give an idea of the proportion of instances that belong to each category
```{r Your Seventeenth Code Chunk}
barplot(table(student_performance_dataset[, 96]), main = names(student_performance_dataset)[96])
```

## Create a Missingness Map to Identify Missing Data
We used a Missingness Map to get an idea of the amount missing data in the dataset. The x-axis of the missingness map shows the attributes of the dataset whereas the y-axis shows the instances in the dataset.Horizontal lines indicate missing data for an instance whereas vertical lines indicate missing data for an attribute. The missingness map requires the “Amelia” package
```{r Your Eightteenth Code Chunk}
if (!is.element("Amelia", installed.packages()[, 1])) {
  install.packages("Amelia", dependencies = TRUE)
}
require("Amelia")

missmap(student_performance_dataset, col = c("red", "grey"), legend = TRUE)
```

## Create a Correlation Plot
We used a Correlation Plot to get an idea of which attributes change together.The function “corrplot()” found in the package “corrplot” is required to perform this. The larger the dot in the correlation plot, the larger the correlation. Blue represents a positive correlation whereas red represents a negative correlation
```{r Your Nineteenth Code Chunk}
if (!is.element("corrplot", installed.packages()[, 1])) {
  install.packages("corrplot", dependencies = TRUE)
}
require("corrplot")
corrplot(cor(student_performance_dataset[, 96:97]), method = "circle")
```

Alternatively, the 'ggcorrplot::ggcorrplot()' function can be used to plot a more visually appealing plot. The code below shows how to install a package in R:
```{r Your twenthieth Code Chunk}
if (!is.element("ggcorrplot", installed.packages()[, 1])) {
  install.packages("ggcorrplot", dependencies = TRUE)
}
require("ggcorrplot")
ggcorrplot(cor(student_performance_dataset[, 96:97]))
```

**etc.** as per the lab submission requirements. Be neat and communicate in a clear and logical manner.
